The Talking Heads Experiment

### 2015-04-22 chap2/chap2.tex
changes to headings and minor mistakes across several chapters
hyphenation, punctuation, positioning, etc for chapters 1-6

\section{The Main Components}

@@

Each of these activities requires knowledge and/or skill
(summarised in the table below). Perceiving requires 
visual processes capable of segmenting images and deriving image 
segments. Conceptualisation requires an ontology, 
a repertoire of perceptually grounded distinctions
that can be applied to a segmented image to yield 
distinctive categories or category combinations that 
may constitute the meaning of the utterance. Verbalising 
a conceptualisation requires 
a {\itshape lexicon} that maps parts of the meaning to words and 
a {\itshape syntax} that specifies how to organise individual words
into a larger complex. 
The hearer must use similar 
knowledge sources in the other direction. He must use 
his lexicon and syntax to reconstruct the meanings expressed
by the utterance, and then use the 
ontology again to apply the meaning to the present 
context to find the referent \tabref{tab:2:findthereferent}\todo{maybe: see Table 2.1??}.

@@

The {\it semantic theory} required to make the 
Talking Heads experiment work is very different from 
the classical textbook approaches to the subject, which
tend to assume that categories can be defined in the 
abstract, independent of their context of use. Grounded
conceptualisation and interpretation processes are necessarily 
strongly situated and context-dependent. Whether `blue 
triangle' is going to be effective in picking out the 
intended topic depends on what else is in
the scene. If all other objects in the 
context are also blue triangles, the game will fail. If there
is only one triangle, it was not really necessary to say that 
it is blue. This situatedness and context-dependence, together with 
the non-verbal hints given by the speaker, are crucial 
keys for restricting the 
possible meanings of a form 
and they therefore help the hearer disambiguate 
utterances or figure out the meaning of an unknown form. But situatedness 
and con\-text-dependence are also major
sources of difficulty, because the same categories may
refer to different things for different agents in different
contexts. Even if both agents agree that a particular form 
names a certain category, 
they may still fail in the game if the category picks
out a different referent, for example because they see 
the context from a slightly different vantage point.

@@

According to this theory, natural language structure is
not the consequence of an 
autonomous innate language device which evolved 
by a genetic mutation or series of mutations. Instead 
cognitive mechanisms and structures already in place have been 
marshalled to get a language system off the ground, even though 
once language became essential for the rapid incorporation
of new individuals and the general organisation of 
activities in human populations, it must have started
to recruit vast amounts of brain capacity and stimulated 
other cognitive faculties, such as categorisation,
episodic memory or problem solving, to become in turn 
much more complex and versatile.\footnote{
There has been an ongoing nature versus nurture debate 
with respect to the origins and acquisition of language. 
The different issues in this debate are already well
illustrated in \cite{Piattelli:1980}.}

>>>>

\section{The main components}

@@

Each of these activities requires knowledge and/or skill
(summarised in the table below). Perceiving requires 
visual processes capable of segmenting images and deriving image 
segments. Conceptualisation requires an ontology, 
a repertoire of perceptually grounded distinctions
that can be applied to a segmented image to yield 
distinctive categories or category combinations that 
may constitute the meaning of the utterance. Verbalising 
a conceptualisation requires 
a {\itshape lexicon} that maps parts of the meaning to words and 
a {\itshape syntax} that specifies how to organise individual words
into a larger complex. 
The hearer must use similar 
knowledge sources in the other direction. He must use 
his lexicon and syntax to reconstruct the meanings expressed
by the utterance, and then use the 
ontology again to apply the meaning to the present 
context to find the referent (see \tabref{tab:2:findthereferent}).\todo{I inserted `(see ...)' here)}

@@

The {\it semantic theory} required to make the 
Talking Heads experiment work is very different from 
the classical textbook approaches to the subject, which
tend to assume that categories can be defined in the 
abstract, independent of their context of use. Grounded
processes of conceptualisation and interpretation are necessarily \todo{Word order changed to fit line.}
strongly situated and context-dependent. Whether `blue 
triangle' is going to be effective in picking out the 
intended topic depends on what else is in
the scene. If all other objects in the 
context are also blue triangles, the game will fail. If there
is only one triangle, it was not really necessary to say that 
it is blue. This situatedness and context-dependence, together with 
the non-verbal hints given by the speaker, are crucial 
keys for restricting the 
possible meanings of a form 
and they therefore help the hearer disambiguate 
utterances or figure out the meaning of an unknown form. But situatedness 
and con\-text-dependence are also major
sources of difficulty, because the same categories may
refer to different things for different agents in different
contexts. Even if both agents agree that a particular form 
names a certain category, 
they may still fail in the game if the category picks
out a different referent, for example because they see 
the context from a slightly different vantage point.

@@

\enlargethispage{1\baselineskip}
According to this theory, natural language structure is
not the consequence of an 
autonomous innate language device which evolved 
by a genetic mutation or series of mutations. Instead 
cognitive mechanisms and structures already in place have been 
marshalled to get a language system off the ground, even though 
once language became essential for the rapid incorporation
of new individuals and the general organisation of 
activities in human populations, it must have started
to recruit vast amounts of brain capacity and stimulated 
other cognitive faculties, such as categorisation,
episodic memory or problem solving, to become in turn 
much more complex and versatile.\footnote{
There has been an ongoing nature versus nurture debate 
with respect to the origins and acquisition of language. 
The different issues in this debate are already well
illustrated in \cite{Piattelli:1980}.}

### 2015-04-22 chap3/chap3.tex
changes to headings and minor mistakes across several chapters
hyphenation, punctuation, positioning, etc for chapters 1-6

For example, suppose you want a robot of the sort seen 
in \figref{f:plate3} to move towards a light source. As originally 
suggested by the cybernetician \cite{Braitenberg:1984}. 
we can do this by mounting two simple light sensors 
to the left and right of the body and by implementing a 
direct dynamical coupling between the output of the light
sensors and the motors. The coupling is such that when
the sensed left light is stronger than the sensed right
light, the motor attached
to the left wheel is slightly decreased and the right
motor slightly increased, so that the robot veers to the left. 
When the right light is stronger, the right motor is 
increased and the left motor decreased, so that the 
robot veers to the right. When these two dynamical 
couplings are put in place together with a
forward movement, a zig-zag behaviour emerges which
brings the robot to the light source. \is{behaviours}

@@

The two way flow of constraints (from perception to 
high level cognitive processing and from high level 
cognitive processing to perception) suggests 
that the brain is best thought of as a massively parallel, 
densely connected processing system, in which multiple 
partial solution structures float up or down, gathering strength 
or weakening with new information entering the system. 
This contrasts with the view that 
information is processed in a serial step-by-step fashion
through tightly compartmentalised modules.\footnote{
See \cite{Fodor:1983}. This book discusses a modular, sequential 
information processing architecture for cognition. A 
non-modular, parallel view 
is sketched in: \cite{Minsky:1985}. 
A more realistic neurophysiological model similar to the 
one underlying the {\it sieve architecture} we have used is
discussed by \cite{Edelman:1987}.} It is true
that we have the illusion that there 
is some {\itshape homunculus}, a little man, which sees a single coherent 
picture of reality and finds quickly the right way to 
verbalise or interpret this picture, but closer examination 
of the actual informational requirements of each subprocess
shows that this can never work. Constraints must flow in all 
directions, because the sensory data has not enough information to allow 
a unique segmentation, and categorisation and language is 
too full of ambiguities to allow a straightforward linear
interpretation.

\section{Sensory Channels}

@@

\subsection{Example channels}
\enlargethispage{1\baselineskip}
The segment characteristics which will be 
used later in various experiments are defined below. 
Their values are all derived by straightforward computations from 
the raw image maps captured by the cameras. 
\begin{itemize} 
\item \textsc{area}: The surface area of a segment
is calculated by simply counting the number of pixels that are part 
of the segment. 
\item \textsc{hpos}, \textsc{vpos}: The x and y values of the 
central-position of a segment. The central position 
is calculated by taking the mid-point of the sides 
of the bounding box. 
\item \textsc{height}: The height of the bounding box. 
\item \textsc{width}: The width of the bounding box. 
\item \textsc{bb-area}: The area of the bounding box,
calculated by multiplying height by width. 
\item \textsc{gray}: The average gray-scale value of the pixels
in a segment. 
\item R, G, B: The average R (redness), G
(greenness), and B (blueness) values in a
segment. To obtain more human-like colour channels, they are
transformed in terms of YB (yellow-blue), RG (red-green), 
BW (black-white), SAT (saturation) and \textsc{brightness} channels. 
\item \textsc{edge-count}: The number of edges in
a segment, for example 3 in the case of 
a triangle. 
\item \textsc{angle-count}: The number of angles, determined on the 
basis of the junctions. 
\item \textsc{ratio}: The ratio between the area of the segment 
and the area of its bounding box, which gives an idea how close 
a figure is to a rectangular shape. 
\end{itemize}

@@

\clearpage
Here are some other examples: 
\begin{enumerate}
\item A set of sensory channels that are sensitive 
to characteristics of moving segments can easily be 
imagined. They include the speed
of movement, the direction of movement (along the 
horizontal position (\textsc{hpos}) and vertical position (\textsc{vpos}) 
dimensions), or the change in area (getting bigger if the object 
approaches or smaller if it moves away). This is the foundation
for categories about spatial change: moving left versus moving right,
approaching versus retracting, speeding up versus slowing down.\footnote{Experiments
using segmentation based on movement
using the dedicated hardware shown in \ref{f:plate6} were carried out by Tony Belpaeme. 
More details can be found in: \cite{Belpaeme:1998} These channels have been used in
language experiments where the agents were 
perceiving and communicating about moving 
images: \cite{Steels:1998}  
The segmentation used in this book 
based on output from the Sony EVI-D31 camera was 
implemented by Danny Van Tieghem. Angus McIntyre integrated
the interfaces to this camera within the \textsc{babel} environment. 
}

@@

It is not at all necessary that the channels operate 
on visual input, they can also use actuator sensors or internal states
like the level of the battery. It is moreover possible to 
consider channels for dynamic states, for 
example by transforming the sensor and actuator
data into state-space representations and 
analysing them in terms of attractors.\footnote{
Several examples of this type of approach are 
found in \cite{Port:1995}}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{  l   l   l   c  } \lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1} & {\itshape saliency}\\ \midrule
\textsc{hpos} & 0.27 & 0.16 & 0.11\\ 
\textsc{vpos} & 0.20 & 0.20 & 0.0\\ 
\textsc{height} & 0.15 & 0.15 & 0.0\\ 
\textsc{width} & 0.10 & 0.11 & 0.01\\ 
\textsc{area} & 0.10 & 0.10 & 0.0\\ 
R & 0.23 & 0.25 & 0.02\\ 
G & 0.32 & 0.34 & 0.02\\ 
B & 0.63 & 0.65 & 0.02\\ 
\lspbottomrule
\end{tabular}
\caption{Perceived values for each segment in \figref{f:plate-10}
\label{tab:t-plate10}}
\end{center}
\end{table}

@@

I will extensively make use of this methodology. 
This chapter has focused on perception, the next chapter focuses
exclusively on
conceptualisation, assuming that there is good segmentation and a 
decent set of sensory channels. Chapter 5 turns to the 
problem of lexicalisation, assuming that 
the agents have a shared repertoire of meanings and agree on 
what meaning to use in a particular game. Chapter 6 then puts the solutions 
for conceptualisation and lexicalisation together by coupling their
inputs and outputs and establishing the appropriate feedback connections. 
Chapter 7 then couples this system to the perceptual processes discussed
in this chapter, so that we get back to our 
original goal: understand how perceptually grounded language communication 
is possible.

@@

\begin{table}[b]
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
 & {\itshape \textsc{hpos}} & {\itshape \textsc{vpos}} & {\itshape \textsc{height}} & {\itshape \textsc{width}} & {\itshape \textsc{gray}} & {\itshape \textsc{ratio}}  & {\itshape \textsc{area}} \\ \midrule
0 & 116 & 166 & 293 & 293 & 0.777 & 1.0 & 85849 \\ 
1 & 692 & 317 & 148 & 224 & 0.449 & 1.0 & 33152 \\ 
2 & 192 & 64 & 137 & 137 & 0.408 & 1.0 & 18769 \\ 
3 & 167 & 770 & 277 & 277 & 0.201 & 1.0 & 76729 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geom}Sensory data for the scene shown in \figref{geom3}.}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
& {\itshape \textsc{hpos}} & {\itshape \textsc{vpos}} & {\itshape \textsc{height}} & {\itshape \textsc{width}} & {\itshape \textsc{gray}} & {\itshape \textsc{ratio}}  & {\itshape \textsc{area}} \\ \midrule
0 & 0.09 & 0.13 &  0.64 &  0.64 &  0.78 &  1.0 & 0.95 \\ 
1 & 0.55 & 0.25 & 0.16 & 0.41 & 0.45 & 1.0 & 0.37 \\ 
2 & 0.15 & 0.05 & 0.12 & 0.12 & 0.41 & 1.0 & 0.21 \\ 
3 & 0.13 & 0.62 & 0.59 & 0.59 & 0.20 & 1.0 & 0.85 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geomscaled}Sensory data after scaling for the scene shown in \figref{geom3}.}
\end{table}

@@

Working with simulations has obvious advantages for 
speeding up development and systematic testing, 
but it does not replace experimentation with 
physical robots. It is true that 
building an experimental apparatus such as we built
for the Talking Heads experiment is a very 
non-trivial engineering project, particularly because 
of the teleporting infrastructure that enables the agents 
to travel to multiple sites and thus 
experience different physical environments or the same 
environment from different points of view. 
One might therefore wonder why we have persisted
to go to such great length in constructing
a real-world physical infrastructure. Why are
experiments with robotic agents necessary or 
desirable? Is it not enough to conduct simulation
experiments given that they can be done so much 
more easily?

>>>>

For example, suppose you want a robot of the sort seen 
in \figref{f:plate3} to move towards a light source. As originally 
suggested by the cybernetician \cite{Braitenberg:1984}, 
we can do this by mounting two simple light sensors 
to the left and right of the body and by implementing a 
direct dynamical coupling between the output of the light
sensors and the motors. The coupling is such that when
the sensed left light is stronger than the sensed right
light, the motor attached
to the left wheel is slightly decreased and the right
motor slightly increased, so that the robot veers to the left. 
When the right light is stronger, the right motor is 
increased and the left motor decreased, so that the 
robot veers to the right. When these two dynamical 
couplings are put in place together with a
forward movement, a zig-zag behaviour emerges which
brings the robot to the light source. \is{behaviours}

@@

The two way flow of constraints (from perception to 
high level cognitive processing and from high level 
cognitive processing to perception) suggests 
that the brain is best thought of as a massively parallel, 
densely connected processing system, in which multiple 
partial solution structures float up or down, gathering strength 
or weakening with new information entering the system. 
This contrasts with the view that \enlargethispage{1\baselineskip}
information is processed in a serial step-by-step fashion
through tightly compartmentalised modules.\footnote{
See \cite{Fodor:1983}. This book discusses a modular, sequential 
information processing architecture for cognition. A 
non-modular, parallel view 
is sketched in: \cite{Minsky:1985}. 
A more realistic neurophysiological model similar to the 
one underlying the {\it sieve architecture} we have used is
discussed by \cite{Edelman:1987}.} It is true
that we have the illusion that there 
is some {\itshape homunculus}, a little man, which sees a single coherent 
picture of reality and finds quickly the right way to 
verbalise or interpret this picture, but closer examination 
of the actual informational requirements of each subprocess
shows that this can never work. Constraints must flow in all 
directions, because the sensory data has not enough information to allow 
a unique segmentation, and categorisation and language is 
too full of ambiguities to allow a straightforward linear
interpretation.

\section{Sensory channels}

@@

\subsection{Example channels}
\enlargethispage{1\baselineskip}
The segment characteristics which will be 
used later in various experiments are defined below. 
Their values are all derived by straightforward computations from 
the raw image maps captured by the cameras. 
\begin{itemize} 
\item \textsc{area}: The surface area of a segment
is calculated by simply counting the number of pixels that are part 
of the segment. 
\item \textsc{hpos}, \textsc{vpos}: The x and y values of the 
central-position of a segment. The central position 
is calculated by taking the mid-point of the sides 
of the bounding box. 
\item \textsc{height}: The height of the bounding box. 
\item \textsc{width}: The width of the bounding box. 
\item \textsc{bb-area}: The area of the bounding box,
calculated by multiplying height by width. 
\item \textsc{gray}: The average gray-scale value of the pixels
in a segment. 
\item \textsc{r}, \textsc{g}, \textsc{b}: The average \textsc{r} (redness), \textsc{g}
(greenness), and \textsc{b} (blueness) values in a
segment. To obtain more human-like colour channels, they are
transformed in terms of \textsc{yb} (yellow-blue), \textsc{rg} (red-green), 
\textsc{bw} (black-white), \textsc{sat} (saturation) and \textsc{brightness} channels. 
\item \textsc{edge-count}: The number of edges in
a segment, for example 3 in the case of 
a triangle. 
\item \textsc{angle-count}: The number of angles, determined on the 
basis of the junctions. 
\item \textsc{ratio}: The ratio between the area of the segment 
and the area of its bounding box, which gives an idea how close 
a figure is to a rectangular shape. 
\end{itemize}

@@

\clearpage
Here are some other examples: 
\begin{enumerate}
\item A set of sensory channels that are sensitive 
to characteristics of moving segments can easily be 
imagined. They include the speed
of movement, the direction of movement (along the 
horizontal position (\textsc{hpos}) and vertical position (\textsc{vpos}) 
dimensions), or the change in area (getting bigger if the object 
approaches or smaller if it moves away). This is the foundation
for categories about spatial change: moving left versus moving right,
approaching versus retracting, speeding up versus slowing down.\footnote{Experiments
using segmentation based on movement
using the dedicated hardware shown in \figref{f:plate6} were carried out by Tony Belpaeme. 
More details can be found in: \cite{Belpaeme:1998} These channels have been used in
language experiments where the agents were 
perceiving and communicating about moving 
images: \cite{Steels:1998}  
The segmentation used in this book 
based on output from the Sony EVI-D31 camera was 
implemented by Danny Van Tieghem. Angus McIntyre integrated
the interfaces to this camera within the \textsc{babel} environment. 
}

@@

It is not at all necessary that the channels operate 
on visual input, they can also use actuator sensors or internal states
like the level of the battery. It is moreover possible to 
consider channels for dynamic states, for 
example by transforming the sensor and actuator
data into state-space representations and 
analysing them in terms of attractors.\footnote{
Several examples of this type of approach are 
found in \cite{Port:1995}.}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{  l   l   l   c  } \lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1} & {\itshape saliency}\\ \midrule
\textsc{hpos} & 0.27 & 0.16 & 0.11\\ 
\textsc{vpos} & 0.20 & 0.20 & 0.0\\ 
\textsc{height} & 0.15 & 0.15 & 0.0\\ 
\textsc{width} & 0.10 & 0.11 & 0.01\\ 
\textsc{area} & 0.10 & 0.10 & 0.0\\ 
\textsc{R} & 0.23 & 0.25 & 0.02\\ 
\textsc{G} & 0.32 & 0.34 & 0.02\\ 
\textsc{B} & 0.63 & 0.65 & 0.02\\ 
\lspbottomrule
\end{tabular}
\caption{Perceived values for each segment in \figref{f:plate-10}
\label{tab:t-plate10}}
\end{center}
\end{table}

@@

I will extensively make use of this methodology. 
This chapter has focused on perception, the next chapter focuses
exclusively on
conceptualisation, assuming that there is good segmentation and a 
decent set of sensory channels. \chapref{chap:5} turns to the 
problem of lexicalisation, assuming that 
the agents have a shared repertoire of meanings and agree on 
what meaning to use in a particular game. \chapref{chap:6} then puts the solutions 
for conceptualisation and lexicalisation together by coupling their
inputs and outputs and establishing the appropriate feedback connections. 
\chapref{chap:7} then couples this system to the perceptual processes discussed
in this chapter, so that we get back to our 
original goal: understand how perceptually grounded language communication 
is possible.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
 & {\itshape \textsc{hpos}} & {\itshape \textsc{vpos}} & {\itshape \textsc{height}} & {\itshape \textsc{width}} & {\itshape \textsc{gray}} & {\itshape \textsc{ratio}}  & {\itshape \textsc{area}} \\ \midrule
0 & 116 & 166 & 293 & 293 & 0.777 & 1.0 & 85849 \\ 
1 & 692 & 317 & 148 & 224 & 0.449 & 1.0 & 33152 \\ 
2 & 192 & 64 & 137 & 137 & 0.408 & 1.0 & 18769 \\ 
3 & 167 & 770 & 277 & 277 & 0.201 & 1.0 & 76729 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geom}Sensory data for the scene shown in \figref{geom3}.}
\end{table}

\begin{table}[h]
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
& {\itshape \textsc{hpos}} & {\itshape \textsc{vpos}} & {\itshape \textsc{height}} & {\itshape \textsc{width}} & {\itshape \textsc{gray}} & {\itshape \textsc{ratio}}  & {\itshape \textsc{area}} \\ \midrule
0 & 0.09 & 0.13 &  0.64 &  0.64 &  0.78 &  1.0 & 0.95 \\ 
1 & 0.55 & 0.25 & 0.16 & 0.41 & 0.45 & 1.0 & 0.37 \\ 
2 & 0.15 & 0.05 & 0.12 & 0.12 & 0.41 & 1.0 & 0.21 \\ 
3 & 0.13 & 0.62 & 0.59 & 0.59 & 0.20 & 1.0 & 0.85 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geomscaled}Sensory data after scaling for the scene shown in \figref{geom3}.}
\end{table}

@@

Working \todo{3.5.1 cannot remain the only subsection of 3.5} with simulations has obvious advantages for 
speeding up development and systematic testing, 
but it does not replace experimentation with 
physical robots. It is true that 
building an experimental apparatus such as we built
for the Talking Heads experiment is a very 
non-trivial engineering project, particularly because 
of the teleporting infrastructure that enables the agents 
to travel to multiple sites and thus 
experience different physical environments or the same 
environment from different points of view. 
One might therefore wonder why we have persisted
to go to such great length in constructing
a real-world physical infrastructure. Why are
experiments with robotic agents necessary or 
desirable? Is it not enough to conduct simulation
experiments given that they can be done so much 
more easily?

### 2015-04-22 chap4/chap4.tex
cleaned up after sanity check
hyphenation, punctuation, positioning, etc for chapters 1-6

The previous chapter laid the perceptual groundwork 
for investigating conceptualisation and language 
communication. It proposed a set of mechanisms for segmenting a 
scene and extracting characteristics about each 
segment. The next task of the agent is to use these results
from the perceptual layer to categorise 
and hence conceptualise the
scene. For the time being, I will 
ignore the full complexity of natural
language meaning and focus only on the most 
simple type of conceptualisation one can
imagine, namely categories, which logicians refer 
to as unary predicates, as well as conjunctive 
combinations of categories. Words like `blue', `light' or
`square' name such categories. We not only need a way 
that the Talking Heads can use such categories to conceptualise
a scene based on visual perception, but we also need 
mechanisms that can explain how such categories might 
develop or be acquired by each agent autonomously and 
without supervised training. This is clearly an enormous
challenge and no known universally accepted solution 
exists to this problem.

Categorisation has fascinated philosophers and scientists
since the onset \todo{beginning was replaced with onset here} of thought, not only because it is one
of the most fundamental capabilities of the human mind, 
but also because the subject matter immediately raises 
some intriguing paradoxes and puzzles. First of all, 
we have already seen that there is an enormous 
gap between the symbolic world of objects and categories
and the subsymbolic world of analog sensori-motor signal streams. 
A particular sensory signal is highly context-dependent and
inherently noisy due to the partial
unreliability of sensors and their limited accuracy. 
Sensory processing, transformation, and scaling 
go some way to achieve perceptual constancy
but cognitive processes must clearly make up for
the fleeting erratic nature of reality. 
Here we are immediately confronted with a first paradox. 
Univocal categorisation seems only possible when 
an interpretation of reality is within reach, for example
when there are strong expectations, possibly coming 
from language utterances. But this interpretation
depends itself on categorisation. How can this apparent
causal circularity be broken?

@@

Then there is a third paradox, first suggested
by Jean-Jacques Rousseau. In order to communicate,
the speaker and the hearer need to share the building
blocks of the conceptualisations underlying their communication. 
If I say "the wine glass on the table", I expect 
the hearer to be able to recognise what objects are 
tables, when something is on the table or not on the table, and 
when something is a wine glass versus another kind 
of glass. But if every individual learns categories independently
and autonomously, how can they ever become shared? It is 
assumed that language helps in establishing a shared ontology within 
a language community but language itself depends on 
shared categories, so how can the whole system ever 
get off the ground; how can this chicken and egg situation
be broken?

@@

Even supposing that there is a strong genetic
determination of micro-level brain structure, there is 
still the question how the hundreds of thousands of concepts 
employed by adult human beings might have become 
included in the human genome. Saying that a perceptually 
grounded category is innate does not explain anything.
One has to show a plausible evolutionary history for 
the categories hypothesised to be innate and prove 
that the hypothesised concept genes can propagate sufficiently 
fast in the human population.\footnote{
See \cite{Worden:1995}. 
This article argues, based on the genetic difference
between humans and other primates, that there are limits 
to genetic transmission of cognitive content, like 
categories or grammars. For discussions on the speed
of gene spreading compared to cultural evolution, 
see \cite{Cavalli:1995}}

@@

In the case of the Darwinian explanation for the 
evolution of species, a solution is an organism 
capable of surviving in a given environment. Types 
of organisms are preserved in the genetic material as 
it is copied from one generation to the next. 
Variations are produced due to errors in gene copying, 
mutations, gene insertion, etc. The feedback comes
from the natural environment. Organisms that do not flourish are
less successful in reproduction so that their
genetic material and hence the organisms 
this material generates are less likely to be preserved. 
Selectionism contrasts with Lamarckian instructionism, which 
claims that the organism transmits its adaptations 
and what it has learned during its life time to 
its offspring. \is{instructionism}

@@

\begin{figure}[b]
  \centerline{\includegraphics[width=.40\textwidth]{chap4/figs/game5.pdf}}
\caption{\label{geom}A computer generated scene 
from the \textsc{geom} world.}
\end{figure}

@@

Balls thus trickle down from the top to the bottom of 
a discrimination tree, like in the Pachinko game or 
a lottery machine. The trickling down process can stop as
soon as a distinctive category is found because finer grained 
distinctions are not necessary.

@@

The sensory channel data for the 
same scene now scaled for context is shown in Table \ref{tab:t-game22scaled}. Such context-scaling 
pulls the data further apart and makes categorisation 
therefore much easier and much more 
stable, but the information on saliency is lost and so 
it is no longer clear which channel is to be preferred
for reasons of saliency.

@@

The next example is taken from a series of discrimination 
games played by physically instantiated agents using 
real world images. The series is discussed more extensively 
in Chapter 7. The agent {\bfshape a2} has 
captured the image shown in \figref{f:plate-10}
(top left) and done the necessary segmentation and 
gathering of sensory characteristics. The resulting sensory
values (after sensor-scaling) for the segments are shown
in Table \tabref{tab:t-real}. Object-0 has been selected as the topic.

>>>>

The \todo{This intro is way to long, please include another \textbackslash section} previous chapter laid the perceptual groundwork 
for investigating conceptualisation and language 
communication. It proposed a set of mechanisms for segmenting a 
scene and extracting characteristics about each 
segment. The next task of the agent is to use these results
from the perceptual layer to categorise 
and hence conceptualise the
scene. For the time being, I will 
ignore the full complexity of natural
language meaning and focus only on the most 
simple type of conceptualisation one can
imagine, namely categories, which logicians refer 
to as unary predicates, as well as conjunctive 
combinations of categories. Words like `blue', `light' or
`square' name such categories. We not only need a way 
that the Talking Heads can use such categories to conceptualise
a scene based on visual perception, but we also need 
mechanisms that can explain how such categories might 
develop or be acquired by each agent autonomously and 
without supervised training. This is clearly an enormous
challenge and no known universally accepted solution 
exists to this problem.

Categorisation has fascinated philosophers and scientists
since the onset \todo{beginning was replaced with onset here to fit the line} of thought, not only because it is one
of the most fundamental capabilities of the human mind, 
but also because the subject matter immediately raises 
some intriguing paradoxes and puzzles. First of all, 
we have already seen that there is an enormous 
gap between the symbolic world of objects and categories
and the subsymbolic world of analog sensori-motor signal streams. 
A particular sensory signal is highly context-dependent and
inherently noisy due to the partial
unreliability of sensors and their limited accuracy. 
Sensory processing, transformation, and scaling 
go some way to achieve perceptual constancy
but cognitive processes must clearly make up for
the fleeting erratic nature of reality. 
Here we are immediately confronted with a first paradox. 
Univocal categorisation seems only possible when 
an interpretation of reality is within reach, for example
when there are strong expectations, possibly coming 
from language utterances. But this interpretation
depends itself on categorisation. How can this apparent
causal circularity be broken?

@@

Then there is a third paradox, first suggested
by Jean-Jacques Rousseau. In order to communicate,
the speaker and the hearer need to share the building
blocks of the conceptualisations underlying their communication. 
If I say ``the wine glass on the table'', I expect 
the hearer to be able to recognise what objects are 
tables, when something is on the table or not on the table, and 
when something is a wine glass versus another kind 
of glass. But if every individual learns categories independently
and autonomously, how can they ever become shared? It is 
assumed that language helps in establishing a shared ontology within 
a language community but language itself depends on 
shared categories, so how can the whole system ever 
get off the ground; how can this chicken and egg situation
be broken?

@@

Even supposing that there is a strong genetic
determination of micro-level brain structure, there is 
still the question how the hundreds of thousands of concepts 
employed by adult human beings might have become 
included in the human genome. Saying that a perceptually 
grounded category is innate does not explain anything.
One has to show a plausible evolutionary history for 
the categories hypothesised to be innate and prove 
that the hypothesised concept genes can propagate sufficiently 
fast in the human population.\footnote{
See \cite{Worden:1995}. 
This article argues, based on the genetic difference
between humans and other primates, that there are limits 
to genetic transmission of cognitive content, like 
categories or grammars. For discussions on the speed
of gene spreading compared to cultural evolution, 
see \cite{Cavalli:1995}.}

@@

In the case of the Darwinian explanation for the 
evolution of species, a solution is an organism 
capable of surviving in a given environment. Types 
of organisms are preserved in the genetic material as 
it is copied from one generation to the next. 
Variations are produced due to errors in gene copying, 
mutations, gene insertion, etc. The feedback comes
from the natural environment. Organisms that do not flourish are
less successful in reproduction so that their
genetic material and hence the organisms 
this material generates are less likely to be preserved. 
Selectionism contrasts with Lamarckian instructionism, which 
claims that the organism transmits its adaptations 
and what it has learned during its life time to \enlargethispage{1\baselineskip}
its offspring. \is{instructionism}

@@

\begin{figure}[bh]
  \centerline{\includegraphics[width=.40\textwidth]{chap4/figs/game5.pdf}}
\caption{\label{geom}A computer generated scene 
from the \textsc{geom} world.}
\end{figure}

@@

Balls thus trickle down from the top to the bottom of \enlargethispage{1\baselineskip}
a discrimination tree, like in the Pachinko game or 
a lottery machine. The trickling down process can stop as
soon as a distinctive category is found because finer grained 
distinctions are not necessary.

@@

The sensory channel data for the 
same scene now scaled for context is shown in \tabref{tab:t-game22scaled}. Such context-scaling 
pulls the data further apart and makes categorisation 
therefore much easier and much more 
stable, but the information on saliency is lost and so 
it is no longer clear which channel is to be preferred
for reasons of saliency.

@@

The next example is taken from a series of discrimination 
games played by physically instantiated agents using 
real world images. The series is discussed more extensively 
in \chapref{chap:7}. The agent {\bfshape a2} has 
captured the image shown in \figref{f:plate-10}
(top left) and done the necessary segmentation and 
gathering of sensory characteristics. The resulting sensory
values (after sensor-scaling) for the segments are shown
in \tabref{tab:t-real}. Object-0 has been selected as the topic.
